---
title: "MuMIn and AICcmodavg"
author: "Amy West"
date: "2022-11-01"
output: html_document
---

# Before we start

## Practice dataset download:
[SD_banding_data](SD_banding_data.csv)

```{r echo=FALSE}
banding<-read.csv("~/Documents/School/Biol7263/MuMIn_bird_data_frame.csv")
```

### About the dataset
This is a subset of the data of the grassland birds I caught and banded in South Dakota. 

Record of bird species caught, where they were caught, and their measurements.

Variables: **species** (using four letter code), **age** (binary, hatch-year/after hatch-year), **mass** (in grams), **wing** (length mm), **tarsus** (length mm), **fat** (score on scale of 1-5), **location**, **taxon**,  **foraging** (how the bird forages), **temp** (temperature in F), **season** ( binary, spring/fall),  **year** (binary 2018, 2019), **time** 


## Packages to download
```
install.packages("MuMIn") 

#Stands for Multi-Model Inference
#Calculates AICc scores
#Automated model generation

install.packages("AICcmodavg") 

#stands for AICc model averaging
#Does other things, but we will primarily use it to make pretty AIC tables
```
## Load libraries
```{r , message=FALSE}
library(MuMIn)
library(AICcmodavg)
```


# Part 1:
# Generalized Linear Models

## r code structure
glm(formula, data, family = gaussian,...)

**Formula:**

follows Y=B0+B1X

in R: dependent variable ~ independent variables

**Data:**

name of data frame

**Family:**

in R: family(link= "link_name")

Options: gaussian, binomial, gamma, poisson

## GLM code example

### Gaussian:
```{r}
#simple way to get rid of NA's
##glm will automatically drop NA's in independent variables, but not dependent 
banding_na<-na.omit(banding)

glm_example<- glm(  #glm function to run model
  mass~  #dependent variables that you want to predict
    species+fat+wing*tarsus, #independent variables that you want to use to predict
                              ## * indicates interaction
  data = banding_na,  #reference data set
  family=gaussian) #reference family name (here is where you would also put your link function)
```


### Binomial:
```{r}
#Trick: turn your binomial variable into 1's and 0's
banding$location_numeric<-ifelse( #name a new variable, and use if else function
  banding$location=="Forb",0,1) #if location equals "Forb" turn into a 0 else turn it into a 1
 
#binomial glm, same as gaussian, but change family name
Band_binomial<- glm(location_numeric~temp, data= banding, family=binomial)
```

### Output model results:
```{r}
#not very useful, gives you intercepts for each variable
glm_example
```

```{r}
#nicer, but this is not a t-test
summary(glm_example)
```

```{r}
#Heck yeah! P-values, but have to go to one of above summaries to get intercept estimates
anova(glm_example, test="F")
```

# Part 2:
# Information Criterion

## Get AIC scores 

In regular R, you can get AIC scores, but not AICc scores :(

Both MuMIn and AICcmodavg have an AICc function
```{r}
AIC(glm_example)
AICc(glm_example)
```

## Get ALL the model combinations
(Warning: may anger statisticians, see step command for step wise model selection)

We want to create an optimal model to predict mass in birds
```{r message=FALSE}
#create a large model with thoughtful variables
band_model<- glm(mass~species+tarsus*wing+fat, data = banding_na, family=gaussian,
                 na.action = na.fail) #have to include line to knit, but it works better if you run code without na.action

#modeling dredging
AICc_band_models <- dredge( #construct all possible models
  band_model,    #use band model as a reference
  rank = "AICc", #use AICc scores to compare
  fixed = "species") #list variables that you want to include in every model
                      ## use NULL if you don't want to include
```

## Get output from model selection

```{r}
#Make a list of all the models
model_list <- get.models(AICc_band_models, #retrieve models from dredged data
                              subset = TRUE) #select which models to retrieve
                                              ##1:10 gives you top 10 models
                                              ## subset = TRUE gives you all the models

model_list[1] #check out info from a single model

#Make a messy but informative AICc table
AICc_table_band<-model.sel(model_list)
```


## Make a nicer AICc Table using AICcmodavg

### Make a list of model names
```{r}
#Trick: create a list of model names
model_name_list<-NULL #make an empty list

for (i in 1:10){
  model_name_list = c(model_name_list, as.character(model_list[[i]][['formula']]))
} #loop through model output to extract formula for each model

model_name_listb <- model_name_list[seq(3, length(model_name_list), 3)] #select every third element from list and put it in a new list
```

### Make a nicer AICc table

```{r}
modavg_table<-aictab(model_list, #make a table with models from your list
                     modnames = model_name_listb, #label the models with your names list
                     second.ord = TRUE,   #Use AICc (FALSE gives you AIC)
                     sort = TRUE) #Order based on model weight

#View table
modavg_table
```


# Part 3:
# Exercise on your own

1) Create a glm model with at least three independent variables using the banding data set or another data set of your choice

2) Determine which independent variables contribute to an optimal model using AICc

3) Create a AICc table using AICctab to show your results
 
 
# Resources:

## Package websites

[MuMIn](https://cran.r-project.org/web/packages/MuMIn/MuMIn.pdf)

[AICcmodavg](https://cran.r-project.org/web/packages/AICcmodavg/AICcmodavg.pdf)

## Info on how to use AIC and model selection
```
?AICtab
# gives 10 suggestions for model selection
```
[Chapter 6 of Rethinking Statistics](https://civil.colorado.edu/~balajir/CVEN6833/bayes-resources/RM-StatRethink-Bayes.pdf)

## Some research papers that use AIC
[Kelly et al. 2008](https://link.springer.com/content/pdf/10.1672/07-152.1.pdf)

[Mazerolle, M. 2006](https://www.researchgate.net/profile/Marc-Mazerolle/publication/233527223_Improving_data_analysis_in_herpetology_using_Akaike's_Information_Criterion_AIC_to_assess_the_strength_of_biological_hypotheses/links/5c9cbeb092851cf0ae9cbfcc/Improving-data-analysis-in-herpetology-using-Akaikes-Information-Criterion-AIC-to-assess-the-strength-of-biological-hypotheses.pdf)

[Robertson et al. 2011](https://journals.plos.org/plosone/article?id=10.1371/journal.pone.0016941)
